---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
editor: 
  markdown: 
    wrap: 72
---

#load dslabs packages and others 
```{r}
library("dslabs") 
library("tidyverse")
library("ggplot2")
```


#look at help file for gapminder data 
```{r}
help(gapminder)
```


#get an overview of data structure 
```{r}
str(gapminder)
```


#get a summary of data 
```{r}
summary(gapminder)
```


#determine the type of object gapminder is class(gapminder)
```{r}
summary(gapminder)
```

#assign African coutnries to their own dataset 
```{r}
africadata <- gapminder %>% filter(continent == "Africa")
```


#get an overview of africadata 
```{r}
str(africadata)
```


#get a summary of africadata 
```{r}
summary(africadata)
```


#assign African countries' infant mortality and life expectancy data to
#its own object 
```{r}
africa_il <- africadata[ , c("infant_mortality","life_expectancy")]
```


#assign African countries' life expectancy and population data to its
#own object 
```{r}
africa_lp <- africadata[ , c("life_expectancy","population")]
```

#get an overview of africa_il and africa_lp 
```{r}
str(africa_il)
str(africa_lp)
```


#get a summary of africa_il and africa_lp 
```{r}
summary(africa_il)
summary(africa_lp)
```


#plot life expectancy as a function of infant mortality
```{r}
ggplot(africa_il, aes(x=infant_mortality, y=life_expectancy)) +
geom_point() 
```

#from the plot there appears to be a negative correlation
#between the two variables

#plot life expectancy as a function of population size 
```{r}
ggplot(africa_lp,aes(x=population, y=life_expectancy)) + geom_point() +
scale_x_continuous(trans="log10") 
```

#from the plot there appears to be a 
#positive correlation between the two variables 
#the "streaks" of data 
#seem to come from the number of years avaiable for each country

#find na values in infant_mortality 
```{r}
subset(africadata,is.na(infant_mortality))
```
 
#there is missing infant_mortality data up
#until 1981 and again in 2016, so we should choose a year like 2000 to
#avoid any NAs

#make a dataset only including data from 2000 
```{r}
africadata_2000 <-africadata %>% filter(year == "2000")
```


#get overview and summary of africadata_2000 
```{r}
str(africadata_2000)
summary(africadata_2000)
```


#plot life expectancy as a function of infant mortality with 2000 data
```{r}
ggplot(africadata_2000, aes(x=infant_mortality, y=life_expectancy)) +
geom_point()
```
 
#from the plot there seems to be a negative correlation
#between the two variables

#plot life expectancy as a function of population size with 2000 data
```{r}
ggplot(africadata_2000, aes(x=population, y=life_expectancy)) +
geom_point() + scale_x_continuous(trans="log10") 
```

#from the plot there
#is no clear correlation for the two variables in this dataset

#fit 2000 data to a linear model with life_expectancy as the outcome and
#infant_mortality as the predictor 
```{r}
fit1 <- lm(life_expectancy ~ infant_mortality, data=africadata_2000)
```


#get a summary of fit1 
```{r}
summary(fit1) 
```

#the p-value of 2.83e-08 show that
#there is a significant correlation between life expectancy and infant
#mortality in the 2000 dataset

#fit 2000 data to a linear model with life_expectancy as the outcome and
#population as the predictor 
```{r}
fit2 <- lm(life_expectancy ~ population, data=africadata_2000)
```


#get a summary of fit2 
```{r}
summary(fit2) 
```

#the p-vlaue of 0.616 shows there
#is not a significant correlation between life expectancy and population
#in the 2000 dataset


# From this line, down is work from Connor Ross :) 

```{r}
## install.package('broom') was previously done

## Load package 'broom'
library(broom)

```



## Fitting GDP against Infant Mortality and Population in a Linear Model

```{r}
## Feeling a little curious so I made some graphs...

### Wonder what effects GDP on IM looks like...

ggplot2::ggplot(data = africadata_2000, mapping = aes(x = log(gdp),
                                                      y = log(infant_mortality))) +
  geom_point() + labs(x = "Log GDP", y = "Log Infant Mortality", 
                      title = "Effects of GDP on Infant Mortality in 2000")


### Cool, cool. Now what if I do this with population...
ggplot2::ggplot(data = africadata_2000, mapping = aes(x = log(gdp),
                                                      y = log(population))) +
  geom_point() + labs(x = "Log GDP", y = "Log Population", 
                      title = "Effects of GDP on Population in 2000")
#### ^^ NICEEEEE!


## Fitting GDP against Infant Mortality
fit3 <- lm(gdp ~ infant_mortality, data = africadata_2000)

### Fit3 results
tidy(fit3)


## Fitting GDP against Population
fit4 <- lm(gdp ~ population, data=africadata_2000)

### Fit 4 results
tidy(fit4)


```

Fit 3 model suggests that there may be a negative association between GDP and 
Infant Mortality. Since the p-value is less than our alpha of 0.05 (p < 0.02),
we find these results to be significant. Fit 3 model suggest that there may be 
a positive association between GDP and population. Since the p-value is less
than our alpha of 0.05 (p < 0.001), we can reject the null hypothesis
(no association).













