Exploration

Loading required packages and cleaned data. 
```{r warning=FALSE}
#load required packages
library(tidyverse)
library(here)
```

```{r}
#load data
flu2 <- readRDS(here("fluanalysis", "data", "flu2.rds"))
```

Summary statistics for BodyTemp and Nausea
```{r}
#provide summary data for important variables
flu2 %>% pull(Nausea) %>% summary()
flu2 %>% pull(BodyTemp) %>% summary()
```
Distribution of BodyTemp
```{r}
#look at the distribution of BodyTemp
flu2 %>% ggplot(aes(x=BodyTemp)) + geom_histogram() 
```
Looks like of the body temperatures cluster around 98.5 degrees with some right skew. 

Lets take a look ath the relationship between BodyTemp and some predictors. 
```{r}
flu2 %>% ggplot(aes(x=Nausea, y=BodyTemp)) + geom_boxplot()
```
Seems like median body temperature is just slightly higher in those with nausea than without. 

```{r}
flu2 %>% ggplot(aes(x=CoughIntensity, y=BodyTemp)) + geom_boxplot()
```
Median body temperature looks to increase as cough intensity increases. 

```{r}
flu2 %>% ggplot(aes(x=Sneeze, y=BodyTemp)) + geom_boxplot()
```
Looks like people who did not report sneezing have a higher median body temperature. Weird. 

```{r}
flu2 %>% ggplot(aes(x=Weakness, y=BodyTemp)) + geom_boxplot()
```
Median body temperature seems to increase as weakness increases. 
