Model Evaluation

1. Loading packages and cleaned data
```{r}
# load packages
library(tidyverse)
library(here)
library(tidymodels)
library(performance)
library(skimr)
```

```{r}
# load cleaned data
flu2 <- readRDS(here("fluanalysis", "data", "flu2.rds"))
```

2. Split data
```{r}
set.seed(222)
# Put 3/4 of the data into the training set 
data_split <- initial_split(flu2, prop = 3/4)

# Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

3. Create recipe
```{r}
flu_rec <- 
  recipe(Nausea ~ ., data = train_data) 
```

4. Fit model with recipe
```{r}
#setting logistic regression engine
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")
```

```{r}
#creating workflow using model and recipe
flu_wflow <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flu_rec)
flu_wflow
```

```{r}
#make function to prepare the recipe and train the model 
flu_fit <- 
  flu_wflow %>% 
  fit(data = train_data)
```

```{r}
#pull fitting model object and make tibble of model coefficients
flu_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

5. Using trained workflow to predict (test data)
```{r}
#use trained model to predict with unseen data
predict(flu_fit, test_data)
```
```{r}
#save above together to use to get ROC
flu_aug <- 
  augment(flu_fit, test_data)
```
```{r}
#view the data
flu_aug %>%
  select(Nausea, .pred_No, .pred_Yes)
```

```{r}
#create ROC curve with predicted class probabilities 
flu_aug %>% 
  roc_curve(truth = Nausea, .pred_No) %>% 
  autoplot()
```

```{r}
#ROC estimate
flu_aug %>% 
  roc_auc(truth = Nausea, .pred_No) 
```

Refitting with the main predicotr (RunnyNose) and using the trained model to predict with test and train data. 

7. Alternative Model
```{r}
flu_rec2 <- 
  recipe(Nausea ~ RunnyNose, data = train_data) 
```


```{r}
flu_wflow2 <- 
  workflow() %>% 
  add_model(lr_mod) %>% 
  add_recipe(flu_rec2)
flu_wflow2
```

```{r}
flu_fit2 <- 
  flu_wflow2 %>% 
  fit(data = train_data)
```

```{r}
flu_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

8. Using trained workflow to predict (test data)
```{r}
predict(flu_fit2, test_data)
```

```{r}
flu_aug3 <- 
  augment(flu_fit2, test_data)
```

```{r}
flu_aug3 %>%
  select(Nausea, .pred_No, .pred_Yes)
```

```{r}
flu_aug3 %>% 
  roc_curve(truth = Nausea, .pred_No) %>% 
  autoplot()
```

```{r}
flu_aug3 %>% 
  roc_auc(truth = Nausea, .pred_No) 
```

The model using only runny nose as a predictor had lower ROC than the model with all predictors. 